---
title: "Barack Obama Retweets Network"
author: Mateusz Zaremba
date: December 16, 2019
output: 
  pdf_document:
    toc: false
    toc_depth: 2
    number_sections: true
    fig_caption: false
    df_print: kable
    highlight: tango
    citation_package: natbib
    keep_tex: false
fontzise: 11pt
geometry: margin=1.5in
documentclass: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE, tidy = FALSE, cache = FALSE)
```

# Methods
## Network size metrics
The data was downloaded from WEBSITE and prepared using R; on the ‘Network Repository’ website we can find already calculated network data statistics, like number of nodes, number of edges, average degree etc.; these are also already calculated at the repository’s website (Rossi and Ahmed, 2015). I created a network graph in python programming language using networkx library and compared the obtained statistics with the originals. Turns out they are exactly the same:

|   | Calculated Statistics | Website Statistics |
| - | -------------------| --------------------- |
| Number of nodes | 9631 | 9.6K |
| Number of edges | 9775 | 9.8K |
| Average degree  | 2.0299 | 2 |
Table: The results are the same - we can be pretty confident we have created our network graph in python correctly.

## Network Structure Metrics
Our graph in undirected so it does not contain in- or out-degrees. Instead, we found the highest degree node - it had an id of 2506 and degree of 7655; this was likely Barack Obama’s twitter account since it is the most connected.

## Network Density
The network’s density is equal to 0.000210789557302036, which is close to zero, meaning our graph is close to being ‘fully disconnected’.

## Shortest Path Between Two Nodes
We will find the shortest path between top two nodes with the highest degree and between the nodes with the highest and the lowest degree.

### Top Two
The top two nodes with the highest degree have ids consequently: 2506 (Barack Obama) and 9302. Shortest path between these two nodes is: ['2506', '8474', '9302']; the length of this path is 2, which means they are not connected directly.

### Highest and Lowest Degree
As already mentioned, the id of the node with the highest degree is: 2506 and the id of the node with the lowest degree is: 2709; the shortest path between these two nodes is: ['2506', '2709'] and its length is 1, meaning the nodes are connected directly.

## Identifying Network Communities
We managed to identify 138 community groups:

`Counter({1: 7414, 6: 611, 7: 158, 12: 114, 4: 106, 2: 102, 50: 84, 5: 69, 88: 66, 61: 61, 127: 53, 35: 49, 19: 39, 25: 38, 17: 27, 111: 27, 77: 26, 23: 25, 112: 25, 0: 22, 49: 22, 26: 20, 54: 15, 37: 14, 102: 14, 22: 13, 41: 12, 48: 12, 55: 12, 9: 11, 15: 10, 32: 10, 75: 10, 97: 10, 34: 9, 38: 9, 82: 9, 85: 9, 13: 8, 3: 7, 29: 7, 84: 7, 105: 7, 108: 7, 68: 6, 78: 6, 83: 6, 117: 6, 86: 5, 110: 5, 8: 4, 27: 4, 31: 4, 62: 4, 67: 4, 79: 4, 93: 4, 100: 4, 129: 4, 10: 3, 20: 3, 36: 3, 40: 3, 47: 3, 66: 3, 70: 3, 80: 3, 81: 3, 87: 3, 90: 3, 95: 3, 104: 3, 107: 3, 114: 3, 125: 3, 133: 3, 134: 3, 139: 3, 11: 2, 14: 2, 16: 2, 18: 2, 21: 2, 24: 2, 28: 2, 30: 2, 33: 2, 39: 2, 42: 2, 43: 2, 44: 2, 45: 2, 46: 2, 51: 2, 52: 2, 53: 2, 56: 2, 57: 2, 58: 2, 59: 2, 60: 2, 63: 2, 64: 2, 65: 2, 69: 2, 71: 2, 72: 2, 73: 2, 74: 2, 76: 2, 89: 2, 91: 2, 92: 2, 94: 2, 96: 2, 98: 2, 99: 2, 101: 2, 103: 2, 106: 2, 109: 2, 113: 2, 115: 2, 116: 2, 118: 2, 119: 2, 120: 2, 121: 2, 122: 2, 123: 2, 124: 2, 126: 2, 128: 2, 130: 2, 131: 2, 132: 2, 135: 2, 136: 2, 137: 2, 138: 2})`

It was calculated that the sizes of the communities range from 2 to 7414. We can again assume that the biggest community is likely to be centred around Barack Obama.

## Network Structure Connectivity
Investigation reveals that the Barack Obama network is fully connected, and it has no sub-components; this is not surprising because the edges are retweets, nodes are twitter users and the network consists only of users who retweeted Barack Obamas posts.

## Network Hubs/Brokers
Betweenness and closeness centrality was successfully calculated (Although, it took almost 40 minutes to compute) and sorted from the highest to the lowest score for top 20 results; but a `(PowerIterationFailedConvergence(...), 'power iteration failed to converge within 100 iterations')`
error kept occurring when calculating node eigenvector centrality; replacing `nx.eigenvector_centrality` with `nx.eigenvector_centrality_numpy` solved the issue (Stack Overflow, 2019) and after the fix, the computation was almost instantaneous.

Barack Obama’s node (id = 2506) had the highest score in closeness, betweenness and eigenvector centrality, which means:

- The highest closeness centrality score - it is the farthest away from all other nodes in the network or – it takes the most time to spread information sequentially from it to other nodes (Sciencedirect.com, 2019).
- The highest betweenness centrality score – it has the highest number of distinct paths that strictly contain it in between (Sci.unich.it, 2019).
- The highest eigenvector centrality score – it is the most influential node in the network.

# Results
It was also assumed that the biggest community of 7414 would be Barack Obama’s first-degree neighbours. This cannot be really well seen on the graph without any colouring or sizing:
```{r, out.width="0.55\\linewidth", keepaspectratio, include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("/Users/mateuszzaremba/dev/R/Networks/Network/vanilla.pdf")
```

But we can immediately see Barack Obama’s node when we distinguish the node’s degree using sizing and colouring:
```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("/Users/mateuszzaremba/dev/R/Networks/Network/size\ colour\ node\ degree.pdf")
```

The graph becomes even more readable when we use colouring to represent the node’s score of closeness and the node’s sizing to represent its measure of betweenness:
```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("/Users/mateuszzaremba/dev/R/Networks/Network/colour\ closeness\ -\ size\ betweenness\ with\ NO\ lgl.pdf")
```

The same graph was plotted using `Large Graph Layout` function:
```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("/Users/mateuszzaremba/dev/R/Networks/Network/colour\ closeness\ -\ size\ betweenness\ with\ lgl.pdf")
```

Interesting results were achieved using `Gephi` (Gephi.org, 2019), were it is clearly visible that there is one central node in the network (Barack Obama).
```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("/Users/mateuszzaremba/dev/R/Networks/pdf/gephi.pdf")
```


# Conclusion
We managed to successfully identify the node which is very likely to be Barack Obama’s twitter account – the nodes id was 2506; find the shortest path between top two nodes with the highest degrees as well as the node with the highest and the lowest degree; identify 138 communities within the network and their range. 

The size of the network proved it quite difficult to work with; calculation of the node betweenness and closeness centrality took approx. 40 minutes. Plotting did not take as much time - it took only a few moments; a big improvement was noticed when the nodes labels were not being displayed. Nonetheless, an attempt was made to speed up the process using `with_lgl(…)` (Rdrr.io, 2019) function for `Large Graph Layout` but no improvement in rendering speed was noticed apart from a slightly better visual representation of communities. 

One of the things that could be done in the future, when analysing even bigger networks, to improve calculation and render times, could be a usage of a GPU for network analysis. (Mathworks.com, 2019); e.g. Geforce GTS 250 has 450 cores working in parallel versus 4 on the CPU (Kajan and Slačka, 2019). 

Overall, the data was well described and of a good quality; the original analysis provided a lot of good insight (Look table 1) and the size of the data was appropriate for a comprehensive analysis.

# References
Gephi.org. (2019). Gephi - The Open Graph Viz Platform. [online] Available at: https://gephi.org/ [Accessed 16 Dec. 2019].

Kajan, S. and Slačka, J. (2019). COMPUTING OF NEURAL NETWORK ON GRAPHICS CARD. BSc. Slovak University of Technology in Bratislava.

Mathworks.com. (2019). Neural Networks with Parallel and GPU Computing- MATLAB & Simulink. [online] Available at: https://www.mathworks.com/help/deeplearning/ug/neural-networks-with-parallel-and-gpu-computing.html;jsessionid=fad53e284f66e53b4d2665c85a87 [Accessed 16 Dec. 2019].

Rdrr.io. (2019). layout_with_lgl: Large Graph Layout in igraph: Network Analysis and Visualization. [online] Available at: https://rdrr.io/cran/igraph/man/layout_with_lgl.html [Accessed 16 Dec. 2019].

Rossi, R. and Ahmed, N. (2015). The Network Data Repository with Interactive Graph Analytics and Visualization. [online] Network Repository. Available at: http://networkrepository.com/rt-barackobama.php [Accessed 16 Dec. 2019].

Sci.unich.it. (2019). Betweenness Centrality. [online] Available at: https://www.sci.unich.it/~francesc/teaching/network/betweeness.html [Accessed 16 Dec. 2019].

Sciencedirect.com. (2019). Closeness Centrality - an overview | ScienceDirect Topics. [online] Available at: https://www.sciencedirect.com/topics/computer-science/closeness-centrality [Accessed 16 Dec. 2019].

Stack Overflow. (2019). Using networkx to calculate eigenvector centrality. [online] Available at: https://stackoverflow.com/questions/43208737/using-networkx-to-calculate-eigenvector-centrality [Accessed 16 Dec. 2019].



















